---
title: "LTIP Analysis"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Data Acquistion

```{r, echo=FALSE}
library(tidyquant)

# Download historical P&G stock data:
getSymbols("PG", from="1991-01-01", to="2021-04-01", auto.assign=TRUE, warnings=FALSE)
# Download historical S&P500 index data:
getSymbols("SPY", from="1991-01-01", to="2021-04-01", auto.assign=TRUE, warnings=FALSE)
```


```{r}
library(dplyr)

# Join PG & SP500 data-sets by date:
joined_xts <- merge( SPY, PG )

# Adjusted Prices used to reflect stock split valuations:
joined_df <- data.frame(date=index(joined_xts), coredata(joined_xts)) %>% 
              filter(!is.na(SPY.Close)) %>% select(date, SPY.Adjusted, PG.Adjusted)

```


```{r}
library(ggplot2)

ggplot() +
  geom_line(data=joined_df, aes(x=date, y=PG.Adjusted), color="blue") +
  #geom_line(data=joined_df, aes(x=date, y=SPY.Adjusted), color="red") +
  ggtitle("P&G Stock Price (Adjusted)")+
  labs(y="Adjusted Price")
```
# Feature Calculation

```{r}
# Summarize to month/year for simplification:
summary_df <- joined_df %>%
                mutate( year_month = format(joined_df$date,"%Y-%m-01") ) %>%
                mutate( year_month = as.Date(.$year_month, format="%Y-%m-%d" ) ) %>%
                group_by( year_month ) %>%
                  summarize( PG.AVG = mean(PG.Adjusted), SPY.AVG = mean(SPY.Adjusted), .groups="drop")
```


```{r}
# Calculate Year Ago Values:
ya_df <- summary_df %>% 
          mutate( PG.YA = lag( PG.AVG, order_by=year_month, n=12 ) ) %>%
          mutate( SPY.YA = lag( SPY.AVG, order_by=year_month, n=12 ) ) %>%
          mutate ( PG.IYA = PG.AVG / PG.YA ) %>%
          mutate ( SPY.IYA = SPY.AVG / SPY.YA ) %>%
          mutate ( REL_GROWTH = PG.IYA / SPY.IYA ) %>%
          select ( year_month, PG.AVG, PG.YA, PG.IYA, REL_GROWTH ) %>%
          filter(!is.na(REL_GROWTH))
```

```{r}
# Add Future simulation values for PG Price:
simulation_df <-  ya_df %>% 
                    mutate( PG.3Y = lead( PG.AVG, order_by=year_month, n=12*3 ) ) %>%
                    mutate( PG.3Y_INDEX = PG.3Y/PG.AVG ) %>%
                    mutate( PG.5Y = lead( PG.AVG, order_by=year_month, n=12*5 ) ) %>%
                    mutate( PG.5Y_INDEX = PG.5Y/PG.AVG ) %>%
                    mutate( PG.10Y = lead( PG.AVG, order_by=year_month, n=12*10 ) ) %>%
                    mutate( PG.10Y_INDEX = PG.10Y/PG.AVG )
```


```{r}
# Define consistent colors for graphing:
colors <- c("PG.IYA" = "black", "PG.3Y_INDEX" = "red", "PG.5Y_INDEX" = "blue", "PG.10Y_INDEX" = "green" )
```

# Outlier Detection


```{r}
library(ggQC)

ggplot(simulation_df, aes(x=year_month, y=PG.IYA)) +
  geom_line( na.rm=TRUE, group=1, color="black") + 
  stat_QC( method="XmR", auto.label=T, label.digits=2, show.1n2.sigma = F ) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  labs(y="P&G Stock Price IYA", title="P&G Stock Price IYA Control Chart (3 Sigma)")
```

```{r}
ggplot(simulation_df, aes(x=year_month)) +
  geom_line( aes(y=PG.3Y_INDEX, color="PG.3Y_INDEX"), na.rm=TRUE, group=1 ) +
  geom_line( aes(y=PG.5Y_INDEX, color="PG.5Y_INDEX"), na.rm=TRUE, group=1 ) +
  geom_line( aes(y=PG.10Y_INDEX, color="PG.10Y_INDEX"), na.rm=TRUE, group=1 ) +
labs(y="Price Index", color="Legend") + 
scale_color_manual(values = colors) 

```
Period before 1996 has abnormally fast growth leading to outlier returns for that period:

```{r}
# Filtered data-set:
filtered_df <- simulation_df %>% 
                filter( year_month >= "1996-01-01" )
                
```



```{r}
n_3Y = nrow ( filtered_df %>% filter( !is.na( PG.3Y_INDEX ) ) )
n_5Y = nrow ( filtered_df %>% filter( !is.na( PG.5Y_INDEX ) ) )
n_10Y = nrow ( filtered_df %>% filter( !is.na( PG.10Y_INDEX ) ) )

```


## Improve normalization with log transformation

```{r}
# Improvement with log values?
filtered_df <- filtered_df %>% 
                mutate( PG.LOG_3Y_INDEX = log(PG.3Y_INDEX) ) %>%
                mutate( PG.LOG_5Y_INDEX = log(PG.5Y_INDEX) ) %>%
                mutate( PG.LOG_10Y_INDEX = log(PG.10Y_INDEX) )

```


```{r}
library (tidyr)

pvt_df <- filtered_df %>%
            select("PG.LOG_3Y_INDEX", "PG.LOG_5Y_INDEX") %>%
            pivot_longer(
              cols=everything(),
              names_to = "PERIOD",
              values_to = "LOG_INDEX"
            )
```


```{r}
library(grid)
require(gridExtra)

g1 <- ggplot(filtered_df) +
  geom_density( aes(x=PG.LOG_3Y_INDEX, color="PG.3Y_INDEX"), na.rm=TRUE, fill="red", alpha=0.05 ) +
  geom_vline(aes(xintercept=mean(PG.LOG_3Y_INDEX, na.rm=TRUE), color="PG.3Y_INDEX"), linetype="dashed", size=1, alpha=0.25) +
  stat_function(
    fun = dnorm, 
    args = list(mean = mean(filtered_df$PG.LOG_3Y_INDEX, na.rm=T), sd = sd(filtered_df$PG.LOG_3Y_INDEX, na.rm=T)), 
    color="red", linetype="dashed" ) +
labs(x="Price Index", color="Legend") + 
scale_color_manual(values = colors) +
    xlim(-0.5, 1.25)
  
g2 <- ggplot(filtered_df) +
  geom_density( aes(x=PG.LOG_5Y_INDEX, color="PG.5Y_INDEX"), na.rm=TRUE, fill="blue", alpha=0.05 ) +
  geom_vline(aes(xintercept=mean(PG.LOG_5Y_INDEX, na.rm=TRUE), color="PG.5Y_INDEX"), linetype="dashed", size=1, alpha=0.25) +
  stat_function(
    fun = dnorm, 
    args = list(mean = mean(filtered_df$PG.LOG_5Y_INDEX, na.rm=T), sd = sd(filtered_df$PG.LOG_5Y_INDEX, na.rm=T)), 
    color="blue", linetype="dashed" ) +
labs(x="Price Index", color="Legend") + 
scale_color_manual(values = colors) +
    xlim(-0.5, 1.25)
  
g3 <- ggplot(filtered_df) +
  geom_density( aes(x=PG.LOG_10Y_INDEX, color="PG.10Y_INDEX"), na.rm=TRUE, fill="green", alpha=0.05 ) +
  geom_vline(aes(xintercept=mean(PG.LOG_10Y_INDEX, na.rm=TRUE)), color="green", linetype="dashed", size=1, alpha=0.25) +
  stat_function(
    fun = dnorm, 
    args = list(mean = mean(filtered_df$PG.LOG_10Y_INDEX, na.rm=T), sd = sd(filtered_df$PG.LOG_10Y_INDEX, na.rm=T)), 
    color="green", linetype="dashed" ) +
labs(x="Price Index", color="Legend") + 
scale_color_manual(values = colors) +
  xlim(-0.5, 1.25)

grid.arrange(g1, g2, g3, nrow=4, top=textGrob("Distribution of Payout (solid) vs. Normal Distribution Per Period Held"))

```


```{r}
qqnorm ( filtered_df$PG.LOG_3Y_INDEX, col="red")
qqline(filtered_df$PG.LOG_3Y_INDEX, col="red" )

qqnorm ( filtered_df$PG.LOG_5Y_INDEX, col="blue")
qqline(filtered_df$PG.LOG_5Y_INDEX, col="blue" )

qqnorm ( filtered_df$PG.LOG_10Y_INDEX, col="green")
qqline(filtered_df$PG.LOG_10Y_INDEX, col="green" )
```




# Explatory Data Analysis

Review hypothesized key-drivers of stock growth:
* Time held
* Performance relative to S&P
* Performance relative to previous year

## 95% confidence interval for each x-bar

```{r}
library (BSDA)

tsum.test( 
  mean.x = mean(filtered_df$PG.LOG_3Y_INDEX, na.rm=T),
  s.x = sd(filtered_df$PG.LOG_3Y_INDEX, na.rm=T),
  n.x = n_3Y,
  mu=0,
  alternative="greater",
  conf.level= 0.95
)

tsum.test( 
  mean.x = mean(filtered_df$PG.LOG_5Y_INDEX, na.rm=T),
  s.x = sd(filtered_df$PG.LOG_5Y_INDEX, na.rm=T),
  n.x = n_5Y,
  mu=0,
  alternative="greater",
  conf.level= 0.95
)

tsum.test( 
  mean.x = mean(filtered_df$PG.LOG_10Y_INDEX, na.rm=T),
  s.x = sd(filtered_df$PG.LOG_10Y_INDEX, na.rm=T),
  n.x = n_10Y,
  mu=0,
  alternative="greater",
  conf.level= 0.95
)


```

## Time Held: What is probability that 3Y, 5Y & 10Y stock price <= 0 (e.g. Index<=1, Log(Index)<=0) [Assuming ]

(See above distributions)

```{r}
p3Y <- pnorm( 0, mean= mean(filtered_df$PG.LOG_3Y_INDEX, na.rm=T), sd = sd(filtered_df$PG.LOG_3Y_INDEX, na.rm=T), lower.tail=TRUE)
p5Y <- pnorm( 0, mean= mean(filtered_df$PG.LOG_5Y_INDEX, na.rm=T), sd = sd(filtered_df$PG.LOG_5Y_INDEX, na.rm=T), lower.tail=TRUE)
p10Y <- pnorm( 0, mean= mean(filtered_df$PG.LOG_10Y_INDEX, na.rm=T), sd = sd(filtered_df$PG.LOG_10Y_INDEX, na.rm=T), lower.tail=TRUE)

(p3Y*100) 
(p5Y*100)
(p10Y*100)
```
## Are sample means greater for 3/5/10 year periods statistically?

```{r}
tsum.test( 
  mean.y = mean(filtered_df$PG.LOG_3Y_INDEX, na.rm=T), s.y = sd(filtered_df$PG.LOG_3Y_INDEX, na.rm=T), n.y = n_3Y,
  mean.x = mean(filtered_df$PG.LOG_5Y_INDEX, na.rm=T), s.x = sd(filtered_df$PG.LOG_5Y_INDEX, na.rm=T), n.x = n_5Y,
  alternative="greater"
)

tsum.test( 
  mean.y = mean(filtered_df$PG.LOG_5Y_INDEX, na.rm=T), s.y = sd(filtered_df$PG.LOG_5Y_INDEX, na.rm=T), n.y = n_5Y,
  mean.x = mean(filtered_df$PG.LOG_10Y_INDEX, na.rm=T), s.x = sd(filtered_df$PG.LOG_10Y_INDEX, na.rm=T), n.x = n_10Y,
  alternative="greater"
)

```

### Is REL_GROWTH a driver of Price Index?

```{r}
require(gridExtra)
g1 <- ggplot(filtered_df, aes(x=REL_GROWTH)) +
  geom_point( aes(y=PG.3Y_INDEX), color="red", na.rm=TRUE, size=1)

g2 <- ggplot(filtered_df, aes(x=REL_GROWTH)) +
  geom_point( aes(y=PG.5Y_INDEX), color="blue", na.rm=TRUE, size=1)

g3 <- ggplot(filtered_df, aes(x=REL_GROWTH)) +
  geom_point( aes(y=PG.10Y_INDEX), color="green", na.rm=TRUE, size=1)

grid.arrange(g1, g2, g3, nrow=3, top=textGrob("P&G Payout vs. Relative Growth"))

```

```{r}

summary (lm ( PG.3Y_INDEX ~ REL_GROWTH, filtered_df ) )
summary (lm ( PG.5Y_INDEX ~ REL_GROWTH, filtered_df ) )
summary (lm ( PG.10Y_INDEX ~ REL_GROWTH, filtered_df ) )

```

This is a useless predictor so will likely be removed from regression consideration.


### Is IYA a driver of Price Index?


```{r}
require(gridExtra)
g1 <- ggplot(filtered_df, aes(x=PG.IYA)) +
  geom_point( aes(y=PG.3Y_INDEX), color="red", na.rm=TRUE, size=1)

g2 <- ggplot(filtered_df, aes(x=PG.IYA)) +
  geom_point( aes(y=PG.5Y_INDEX), color="blue", na.rm=TRUE, size=1)

g3 <- ggplot(filtered_df, aes(x=PG.IYA)) +
  geom_point( aes(y=PG.10Y_INDEX), color="green", na.rm=TRUE, size=1)

grid.arrange(g1, g2, g3, nrow=4, top=textGrob("P&G Stock Growth vs. "))

```
```{r}

summary (lm ( PG.3Y_INDEX ~ PG.IYA, filtered_df ) )
summary (lm ( PG.5Y_INDEX ~ PG.IYA, filtered_df ) )
summary (lm ( PG.10Y_INDEX ~ PG.IYA, filtered_df ) )

```
# Generate SO_ALLOCATION Payout Data

```{r}
scenarios <- c(0, .25, .50, .75, 1.0)
scenario_df <- filtered_df %>% full_join( data.frame(SO_ALLOCATION=scenarios), by=character() )

so_to_rsu = 3

so_payout <- function ( so_alloc, price_start, price_end ) {
  price_delta <- price_end - price_start
  
  if ( is.na(price_end) ) {
    return(NA)
  }
  
  if ( price_delta < 0 ) {
    return(0)
  }
  return( so_alloc * so_to_rsu * price_delta )
}

rsu_payout <- function (so_alloc, price_end ) {
  return ( (1-so_alloc) * price_end )
}

scenario_df <- scenario_df %>%
                mutate ( PAYOUT.3Y = mapply( function(x,y,z) (so_payout(x,y,z) + rsu_payout(x,z))/y , SO_ALLOCATION, PG.AVG, PG.3Y ) ) %>%
                mutate ( PAYOUT.5Y = mapply( function(x,y,z) (so_payout(x,y,z) + rsu_payout(x,z))/y , SO_ALLOCATION, PG.AVG, PG.5Y ) ) %>%
                mutate ( PAYOUT.10Y = mapply( function(x,y,z) (so_payout(x,y,z) + rsu_payout(x,z))/y , SO_ALLOCATION, PG.AVG, PG.10Y ) ) 

                

```

```{r}
scenario_df$SO_ALLOCATION_CTG <- as.factor(scenario_df$SO_ALLOCATION)

ggplot( scenario_df, aes(x=SO_ALLOCATION_CTG, y=PAYOUT.3Y, color=SO_ALLOCATION_CTG))+
  geom_boxplot(na.rm=T) +
  labs(title="3 Year Stock Optional Allocation Payouts")

ggplot( scenario_df, aes(x=SO_ALLOCATION_CTG, y=PAYOUT.5Y, color=SO_ALLOCATION_CTG))+
  geom_boxplot(na.rm=T)+
  labs(title="5 Year Stock Optional Allocation Payouts")

ggplot( scenario_df, aes(x=SO_ALLOCATION_CTG, y=PAYOUT.10Y, color=SO_ALLOCATION_CTG))+
  geom_boxplot(na.rm = T) +
  labs(title="10 Year Stock Optional Allocation Payouts")


```

```{r}
c_df <- scenario_df %>% filter( SO_ALLOCATION == 1 ) %>% filter(!is.na(PAYOUT.3Y)) %>% select(PAYOUT.3Y)
mean_10_0 <- mean( c_df$PAYOUT.3Y, na.rm = T )
sd_10_0 <- sd( c_df$PAYOUT.3Y, na.rm = T )
nrow_10_0 <- nrow( c_df )

c_df <- scenario_df %>% filter( SO_ALLOCATION == 0 ) %>% select(PAYOUT.3Y)
mean_10_1 <- mean( c_df$PAYOUT.3Y, na.rm = T )
sd_10_1 <- sd( c_df$PAYOUT.3Y, na.rm = T )
nrow_10_1 <- nrow( c_df )

tsum.test( 
  mean.y = mean_10_0, s.y = sd_10_0, n.y = nrow_10_0,
  mean.x = mean_10_1, s.x = sd_10_1, n.x = nrow_10_1,
  alternative="greater"
)


```


# Build Regression Model

```{r}
pvt_df <- scenario_df %>%
          select ( SO_ALLOCATION, REL_GROWTH, PG.IYA, PG.AVG, PAYOUT.3Y, PAYOUT.5Y, PAYOUT.10Y) %>%
          pivot_longer( cols=starts_with("PAYOUT"), names_to="PERIOD_HELD", values_to="PAYOUT" ) %>%
          filter(!is.na(PAYOUT))

pvt_df$SO_ALLOCATION <- as.factor(pvt_df$SO_ALLOCATION)

chart.Correlation( pvt_df %>% select_if(is.numeric) )
```
## Linear Regression

```{r}
train_pcnt=.7
test_pcnt = .3

ind <- sample(2, nrow(pvt_df), replace = T, prob = c(train_pcnt, test_pcnt))


train_data <- pvt_df[ind == 1,]
test_data <- pvt_df[ind == 2,]

linear_model = lm ( PAYOUT ~ SO_ALLOCATION + PG.IYA + PERIOD_HELD, data=train_data )
summary(linear_model)

```

```{r}
library(olsrr)
library(MASS)

(all <- ols_step_all_possible(linear_model)) 

m_opt <- stepAIC(linear_model, direction = "both", trace = F)
summary(m_opt)
```


```{r}
plot ( linear_model )
```

```{r}
library(MLmetrics)
test_data$prediction <- predict ( linear_model, test_data)

# Can't use MAPE since y_true can be zero:
MAE( y_pred=test_data$prediction, y_true=test_data$PAYOUT )

plot ( test_data$PAYOUT, test_data$prediction )

```



```{r}
library(rpart)
library(rpart.plot)

tree <- rpart ( formula = PAYOUT ~ SO_ALLOCATION + REL_GROWTH + PG.IYA + PERIOD_HELD, data=train_data)

rpart.plot(tree)
printcp(tree)
plotcp(tree)
rpart.rules(tree)
print(tree)
summary(tree)
```
```{r}
test_data$tree_prediction = predict( tree, test_data )

# Can't use MAPE since y_true can be zero:
MAE( y_pred=test_data$tree_prediction, y_true=test_data$PAYOUT )

plot ( test_data$PAYOUT, test_data$tree_prediction, main="Tree predictions actual vs. predicted", xlab="Actual Payout", ylab="Predicted Payout")
```

```{r}
library(caret)

cvcontrol <- trainControl(method="repeatedcv", 
                          number = 5,
                          repeats = 2,
                          allowParallel=TRUE)

forest <- train(PAYOUT ~ ., 
                data=train_data,
                method="rf",
                trControl=cvcontrol,
                importance=TRUE)
```



```{r}
library(caret)

plot(varImp(forest))

test_data$forest_prediction <-  predict(forest,  test_data)
MAE( test_data$PAYOUT, test_data$forest_prediction)

# sqrt(mean((test$medv - rf)^2))
# cor(test$medv, rf) ^2

plot(test_data$PAYOUT, test_data$forest_prediction)

```


```{r}
library(lime)

explainer <- lime(train_data[1:5,], forest, n_bins = 5)
explanation <- explain( x = train_data[1:5,], 
                        explainer = explainer, 
                        n_features = 5)
plot_features(explanation)
plot_explanations(explanation)

```

# Follow up question - would it be better to invest externally?



```{r}
ggplot(filtered_df, aes(y=REL_GROWTH)) +
  geom_boxplot(na.rm = T) + 
  labs (title="P&G stock price growth vs. S&P500 growth")

mean(filtered_df$REL_GROWTH)
```



